---
title: "Calendar variables"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(lubridate)

theme_set(theme_minimal())
default_color <- "royal blue"
ozone_label <- expression(paste("Ozone (", mu, "g/", m^3, ")"))

source("utils.R")

tab_model <- read_rds("../../data/data_model.rds")
```

## Summary

Ozone = daily ozone average, between 12 p.m. and 4 p.m.

```{r}
tab_model %>% 
  group_by(stationname, year) %>% 
  summarise(
    `Number of obs` = sum(!is.na(o3_mass_conc)),
    `Number of missing obs` = sum(is.na(o3_mass_conc)),
    `% of missing obs` = scales::percent(
      mean(is.na(o3_mass_conc)), accuracy = 0.1
    )
  ) %>% 
  format_table()
```


## By year

Annual ozone average. The dashed line equals the highest average.

```{r}
tab_model %>% 
  group_by(year, stationname) %>% 
  summarise(
    o3_mass_conc = mean(o3_mass_conc, na.rm = TRUE)
  ) %>% 
  group_by(stationname) %>% 
  mutate(o3_max = max(o3_mass_conc)) %>% 
  ggplot(aes(x = year, y = o3_mass_conc)) +
  geom_line() +
  facet_wrap(vars(stationname), ncol = 3) +
  geom_hline(aes(yintercept = o3_max), linetype = 2, color = default_color) +
  labs(y = ozone_label, x = "Year")
```

Annual ozone average. Error bars built with +/- 2 std. deviation.

```{r}
tab_model %>% 
  group_by(year, stationname) %>% 
  summarise(
    o3_avg = mean(o3_mass_conc, na.rm = TRUE),
    o3_sd = sd(o3_mass_conc, na.rm = TRUE),
    o3_inf = o3_avg - 2 * o3_sd,
    o3_sup = o3_avg + 2 * o3_sd
  ) %>%
  ggplot(aes(x = year, y = o3_avg, color = stationname)) +
  geom_point(show.legend = FALSE) +
  geom_errorbar(
    aes(ymin = o3_inf, ymax = o3_sup), 
    width = 0.2,
    show.legend = FALSE
  ) +
  facet_wrap(vars(stationname), ncol = 3) +
  labs(y = ozone_label, x = "Year")
```

Annual ozone distribution.

```{r}
tab_model %>% 
  mutate(year = as.character(year)) %>% 
  ggplot(aes(x = year, y = o3_mass_conc)) +
  geom_boxplot() +
  facet_wrap(vars(stationname), ncol = 3) +
  labs(y = ozone_label, x = "Year")
```

## By month, all years

Monthly ozone average. The dashed line equals the highest average.

```{r}
tab_model %>% 
  group_by(month, stationname) %>% 
  summarise(
    o3_mass_conc = mean(o3_mass_conc, na.rm = TRUE)
  ) %>% 
  group_by(stationname) %>% 
  mutate(o3_max = max(o3_mass_conc)) %>% 
  ggplot(aes(x = month, y = o3_mass_conc)) +
  geom_line() +
  facet_wrap(vars(stationname), ncol = 3) +
  geom_hline(aes(yintercept = o3_max), linetype = 2, color = default_color) +
  labs(y = ozone_label, x = "Month")
```

Monthly ozone distribution.

```{r}
tab_model %>% 
  mutate(month = str_pad(month, 2, "left", "0")) %>% 
  ggplot(aes(x = month, y = o3_mass_conc)) +
  geom_boxplot() +
  facet_wrap(vars(stationname), ncol = 2) +
  labs(y = ozone_label, x = "Month")
```

## By week of the year, all years

Ozone average by week of the year. The dashed line equals the highest average.

```{r}
tab_model %>% 
  group_by(week, stationname) %>% 
  summarise(
    o3_mass_conc = mean(o3_mass_conc, na.rm = TRUE)
  ) %>% 
  group_by(stationname) %>% 
  mutate(o3_max = max(o3_mass_conc)) %>% 
  ggplot(aes(x = week, y = o3_mass_conc)) +
  geom_line() +
  facet_wrap(vars(stationname), ncol = 3) +
  geom_hline(aes(yintercept = o3_max), linetype = 2, color = default_color) +
  labs(y = ozone_label, x = "Week of the year")
```

## By day of week, all years

Ozone average by day of the week, excluding public holidays.

```{r, fig.height=10, fig.width=8}
tab_model %>% 
  filter(dv_weekday_regular == 1 | dayofweek %in% c(0, 6)) %>%
  mutate(dayofweek = wday(dayofweek + 1, label = TRUE)) %>% 
  group_by(dayofweek, stationname) %>% 
  summarise(
    o3_mass_conc = mean(o3_mass_conc, na.rm = TRUE)
  ) %>% 
  group_by(stationname) %>% 
  mutate(o3_max = max(o3_mass_conc)) %>% 
  ggplot(aes(x = dayofweek, y = o3_mass_conc)) +
  geom_line(aes(group = 1)) +
  geom_label(aes(label = round(o3_mass_conc))) +
  facet_wrap(vars(stationname), ncol = 2) +
  labs(y = ozone_label, x = "Day of week")
```

Public holidays vs regular days

```{r}
tab_model %>% 
  group_by(dv_publicholiday) %>% 
  summarise(
    `Ozone average` = mean(o3_mass_conc, na.rm = TRUE),
    `Number of obs` = sum(!is.na(o3_mass_conc)),
    `Number of missing obs` = sum(is.na(o3_mass_conc)),
    `Prop of missing obs` = scales::percent(
      mean(!is.na(o3_mass_conc)), accuracy = 0.1
    )
  ) %>% 
  mutate(dv_publicholiday = ifelse(dv_publicholiday == 0, "No", "Yes")) %>% 
  rename(`Public holiday` = dv_publicholiday)
```

Ozone average by day of the week for regular days and public holidays.

```{r}
tab_model %>% 
  mutate(
    dayofweek = wday(dayofweek + 1, label = TRUE),
    dv_publicholiday = ifelse(dv_publicholiday == 0, "No", "Yes")
  ) %>% 
  group_by(dayofweek, stationname, dv_publicholiday) %>% 
  summarise(
    o3_mass_conc = mean(o3_mass_conc, na.rm = TRUE)
  ) %>% 
  group_by(stationname) %>% 
  mutate(o3_max = max(o3_mass_conc)) %>% 
  ggplot(aes(x = dayofweek, y = o3_mass_conc, color = dv_publicholiday)) +
  geom_line(aes(group = dv_publicholiday)) +
  facet_wrap(vars(stationname), ncol = 2) +
  labs(y = ozone_label, x = "Day of week", color = "Public holiday")
```

